---
title: "Data Preprocessing Pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Preprocessing Pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

This is an example data preprocessing pipeline. We demonstrate how to preprocess source data in order to use it in modeling part.

```{r}
library(bipolarPreprocessing)
library(dplyr)
```


### Get sample data

```{r}
visits <- get_sample_visits()
mobile_recordings <- get_sample_mobile_recordings()
mobile_chunks <- get_sample_mobile_chunks()
```

### Remove duplicates if exists

Check mobile recordings dataset by name (audio file):
```{r}
recordings_duplicates <- mobile_recordings %>%
  count(name) %>% 
  arrange(-n) %>% 
  filter(n > 1) %>% 
  pull(name)

```

### Dates

Data collected before 2018 is unreliable due to a bug in the library that was used to collect all the parameters of a call. We'll work only with data from 2018.

```{r}
visits <- visits %>%
  filter(visit_date >= "2018-01-01")
mobile_recordings <- mobile_recordings %>%
  filter(create_date >= "2018-01-01")
```
We don't filter `mobile_chunks` as it's not clear when the data was recorded. Instead, we'll use corresponding rows from `mobile_recordings` dataset to get only valid rows by joining the datasets together.


### Prepare datasets for further analysis

Recordings with more than 10 chunks are unreliable.
```{r}
mobile_recordings_filtered <- mobile_recordings %>% 
  filter(chunks_count <= 10)

```

## Calculate aggregates

```{r}
sample_call_parameters <- get_sample_call_parameters()

stats <- tryCatch({
  mobile_chunks %>% 
    tidyr::nest(data = -c(dw_mobilerecording_id, dw_patient_id)) %>% 
    mutate(stats = lapply(data, function(i) {
      i %>%
        select(all_of(sample_call_parameters)) %>% 
        summarise_at(vars(sample_call_parameters), list(mean = mean, sd = sd))
    }))
})

stats <- stats %>% 
  select(dw_mobilerecording_id, dw_patient_id , stats) %>% 
  tidyr::unnest(cols = c(stats))  %>% 
  select(dw_mobilerecording_id,dw_patient_id, ends_with("mean"), ends_with("sd"))

stats
```


## Final dataset preperation 

### Connecting mobile recordings with mobilerchunks
```{r}
callings <- mobile_recordings_filtered %>%
  left_join(stats , by = c("dw_mobilerecording_id","dw_patient_id"))  
  
callings <- callings %>% 
 select( dw_mobilerecording_id, patient_id,create_date, ends_with("mean"), ends_with("sd") ) %>%
  mutate(date = create_date)

       
 
callings #Now we have: dw_mobilerecording_id patient_id create_date and aggregated values for all acoustic features
```


### Joining with visits

```{r}
set.seed(123)
library(bipolarPreprocessing)
visits <- get_sample_visits()
visits$visit_date <- as.Date(visits$visit_date)

#Setting LABEL
visits <-  transform_label_custom(visits , 'hamd_ymrs')

auto_config <- auto_create_phases_config(visits, hamd_ymrs)
auto_config
```


```{r}

# examples
auto_config %>% add_confidence(values = .5)
auto_config %>% add_confidence(values = .5, normalize = FALSE)
auto_config %>% add_confidence(values = c(0.5, 0.5, 0.7, 0.7, 1, 1, 1,1,1,1))
gauss <- auto_config %>% add_confidence(func = dnorm)
plot(gauss$time_point, gauss$confidence, type = "b")
```


### Adding confidence path

```{r}
#insert new rows for each record
config_with_confidence <- auto_config %>% add_confidence(func = dnorm)
extended_visits <- expand_ground_truth_period(
  d = visits,
  config = auto_config,
  phases_col = hamd_ymrs,
  visit_date_col = visit_date,
  patient_id_col = patient_id
)
extended_visits

```


```{r}

extended_visits_clean <- transform_overlapping_phases(extended_visits, config_with_confidence)
extended_visits_clean 

# The next step is merging this data with recordings by 'date' = 'recording_date' (day only)
## First, transform visits with expanded time window into wider form
visits_wide <- extended_visits_clean %>%
  mutate(val = 1) %>%
  tidyr::pivot_wider(names_from = phase, values_from = val, values_fill = 0)


## now join with recordings
model_df <- left_join(
  callings,
  visits_wide,
  by = c("date", "patient_id")
) %>% arrange(patient_id, date, visit_date)

## preview the data:
model_df %>% filter(!is.na(visit_date)) %>% print(n = 25)
```
