---
title: "Modeling with SSFCMeans Pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modeling with SSFCMeans Pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Prepare environment

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
set.seed(123)

library(dplyr)
library(bipolar)
```

Install package `ssfclust` that implements Semi-Supervised Fuzzy C-Means model.

```{r}
# devtools::install_github("ITPsychiatry/ssfclust@refactor")
library(ssfclust)
```

# Read data

The actual files are not kept in this repository - *csv* files are git-ignored.
One must upload the data files on their own.
Next versions of this vignette will contain exemplary data.

```{r}
df <- get_sample_aggregated_data() %>% filter(parametr == "srednia") %>% na.omit()

raw_visits <- get_all_visits() %>% select(patient_id, visit_date, hamd_ymrs)
```

## Structure of `df` dataframe

The `df` dataframe contains summaries of sensors data from phone calls.
One row is a summary of sensors data from a single phone call.

Any `df` dataframe should contain at least:

* composite key `c("patient_id", "create_date")`, where `create_date`
column holds date (YYYY-MM-DD) of the phone call recording,

* summaries of variables collected from each phone call $X$ (these can
be e.g., *mean* of a given variable -- mean jittering for example -- or
standard deviation of a given variable).

The `df` dataframe can, but does not have to, contain a unique identifier
of each phone call.
In the example above, this identifier is kept in the `recording` column.
Note that this is not necessary as **we treat each row as summarising a specific phone call by design**.
It is the responsibility of the user to satisfy that condition for there are no automatic quality checks.

## Structure of `raw_visits` dataframe

Dataframe `visits` should contain:

* composite key `c("patient_id", "visit_date")`,
* information necessary to derive a **label** for each visit;
it can be preprocessed information already in form of labels (e.g., classification
*healthy* or *mania*) or it can be data that still needs postprocessing (e.g., raw data from questionnaires).

In the latter case, postprocessing needs to occur (we point in this vignette where it should occur).

# Data preparation

## dates data type

Cast data type of visit date, get rid of missing observations.

```{r}
raw_visits$visit_date <- as.Date(raw_visits$visit_date)
raw_visits <- raw_visits[-which(is.na(raw_visits$visit_date)), ]
```

## select significant predictors only

Dataframe `sensors_data` that we create here should contain only:

* aforementioned composite key `c("patient_id", "date")` (we rename `create_date` to shorter name `date` for simplicity),
* those sensors' data variables from `df` that are deemed significant variables - they should be encoded in the character vector `x_variables`.

```{r}
x_variables <- c("f0_sma", "f0env_sma", "jitterlocal_sma", "pcm_rmsenergy_sma")

sensors_data <- df %>% 
  select(patient_id, create_date, all_of(x_variables)) %>%
  mutate(date = as.Date(create_date))
```

```{r}
dim(sensors_data)
```

We have `r nrow(sensors_data)` unsupervised observations.
We now need to apply ground truth extrapolation technique.

## prepare `visits` dataframe

The `visits` dataset that we now create is a clean form of `raw_visits`.
It should contain a column $Y = (y_j)^T,\:j=1,\ldots,N,$ containing categorical labels $y_j \in \{y_1, \ldots, y_c\}$.
For example, a following encoding is applied in this vignette: $y_1 = \text{euthymia}$, $y_2 = \text{depression}$, $y_3 = \text{mania}$, $y_4 = \text{mixed}$.

If your `raw_visits` datasets needs postprocessing (e.g. summarising data from questionnaires) to derive such
categorical labels, **this is the place to apply it**.

Our exemplary `visits` dataframe already contains categorical labels in column `hamd_ymrs`:
```{r}
unique(raw_visits$hamd_ymrs)
```

Your final `visits` dataframe should contain only these columns:

* `patient_id`: part of compiste key,
* `visit_date`: date of visit in YYYY-MM-DD format,
* $Y$: column encoding the categorical label assigned to each visit.
You can use your own name, e.g. `label`.

We rename `hamd_ymrs` to `label` in this vignette and create the `visits` dataframe.

```{r}
visits <- transform_label_custom(raw_visits, "hamd_ymrs")
```

# Apply ground truth period extrapolation to data from `visits`

## time windows config preparation for phases

Use the `auto_create_phases_config` function to define time windows for ground truth period extrapolation for each phase separately.

```{r}
# rename name of the function temporarily
auto_create_phases_tw_config <- auto_create_phases_config

auto_tw_config <- auto_create_phases_tw_config(visits, label)
auto_tw_config
```

The `auto_tw_config` we created applied the same time window $[-7, 2]$ for each of the four phases found in the `visits$label` column.

## Label confidence weighing 

We can *weigh* the confidence of label -- expressed as a fraction taking values in
the unit interval $[0, 1]$ -- following the rule: the further from the visit, the smaller the weight (= the smaller the confidence).

Note that by default confidence of each label is equal to $1$.

For example, we can use Gaussian curve (mean=0, sd=5) to apply confidence weighing.

```{r}
tw_config_with_confidence <- auto_tw_config %>%
  add_confidence(func = dnorm, mean=0, sd=5)
```

Let us visualize the confidence extrapolation for `depression` phase, note that
the same strategy applies for the remaining phases (in this vignette).

```{r, fig.dim = c(6, 4)}
with(tw_config_with_confidence %>% filter(phase == "depression"),
     plot(x=time_point, y=confidence, type="b",
          xlab="Day relative to the visit",
          ylab="Label confidence",
          main=paste0(
            "Ground truth period extrapolation for \'depression\' phase\n",
            "Weighing with Normal(0, 5) distribution"),
          ylim=c(0, 1)
     ))
```

## Extrapolate ground truth period onto actual data

We now need to apply *ground truth period extrapolation* configured above
on the actual data collected in the `visits` dataframe.

```{r}
visits_extrapolated <- expand_ground_truth_period(
  d = visits,
  config=auto_tw_config,
  phases_col = label,
  visit_date_col = visit_date,
  patient_id_col = patient_id
)
```

```{r}
visits_extrapolated
```

```{r}
visits_extrapolated_clean <- transform_overlapping_phases(visits_extrapolated, tw_config_with_confidence)

visits_extrapolated_clean
```
Wide format is required for modeling:

```{r}
visits_wide <- visits_extrapolated_clean %>%
  mutate(val = 1) %>%
  tidyr::pivot_wider(
    names_from = phase, 
    values_from = val, 
    values_fill = 0, 
    names_prefix="phase_")
```

# Annotate the data

```{r}
model_df <- left_join(
  sensors_data,
  visits_wide,
  by = c( "patient_id", "date")
) 
```

How much data we have annotated?

```{r}
paste(
  "We annotated ",
  ifelse(is.na(model_df$confidence), 0, 1) |> mean() |> round(2),
  " of the data")
```

# Prepare data for modeling

```{r}
X <- model_df %>% 
  select(all_of(x_variables)) %>%
  as.matrix()
```

```{r}
F_ <- model_df %>% 
  select(starts_with("phase_")) %>%
  replace(is.na(.), 0) %>%
  as.matrix()
```

```{r}
model <- ssfclust::SSFCM(X, C=4)
model_ss <- ssfclust::SSFCM(X, C=4, F_=F_, alpha=1)
```





