---
title: "Example preprocessing steps with mobile recording data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example preprocessing steps with mobile recording data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE, message = FALSE, warning = FALSE
)
```


```{r}
library(bipolarPreprocessing)
library(dplyr)
```


### Get sample data

```{r}
visits <- get_sample_visits()
mobile_recordings <- get_sample_mobile_recordings()
mobile_chunks <- get_sample_mobile_chunks()
```

### What about duplicates?

Check mobile recordings dataset by name (audio file):
```{r}
recordings_duplicates <- mobile_recordings %>%
  count(name) %>% 
  arrange(-n) %>% 
  filter(n > 1) %>% 
  pull(name)

mobile_recordings %>% filter(name %in% recordings_duplicates)
```

Check chunks by chunk id and frame no:

```{r}
chunks_duplicates <- mobile_chunks %>% 
  count(dw_mobilerecordingchunk_id, frame_nr) %>% 
  arrange(-n) %>% 
  filter(n > 1)

if (nrow(chunks_duplicates) > 0) {
  mobile_chunks %>%
    inner_join(chunks_duplicates) %>%
    arrange(dw_mobilerecordingchunk_id, frame_nr)
}
```

No duplicates in our data. If there were any we should filter them out.


### Dates

Data collected before 2018 is unreliable due to a bug in the library that was used to collect all the parameters of a call. We'll work only with data from 2018.

```{r}
visits <- visits %>% filter(visit_date >= "2018-01-01")
mobile_recordings <- mobile_recordings %>% filter(create_date >= "2018-01-01")
```

We don't filter `mobile_chunks` as it's not clear when the data was recorded. Instead, we'll use corresponding rows from `mobile_recordings` dataset to get only valid rows by joining the datasets together.


### Prepare datasets for further analysis

See the distribution of `mobile_recordings$chunks_count` values:

```{r}
mobile_recordings$chunks_count %>% summary()
```

Recordings with more than 10 chunks are unreliable.

```{r}
mobile_recordings_filtered <- mobile_recordings %>% filter(chunks_count <= 10)

unreliable_rec_id <- mobile_recordings %>% 
  filter(chunks_count > 10) %>% 
  select(dw_mobilerecording_id, chunks_count)

unreliable_rec_id
```

## Calculate aggregates

```{r}
sample_call_parameters <- get_sample_call_parameters()

stats <- tryCatch({
  mobile_chunks %>% 
    tidyr::nest(data = -dw_mobilerecording_id) %>% 
    mutate(stats = lapply(data, function(i) {
      i %>%
        select(all_of(sample_call_parameters)) %>% 
        summarise_at(vars(sample_call_parameters), list(mean = mean, sd = sd))
    }))
})

stats <- stats %>% 
  select(dw_mobilerecording_id, stats) %>% 
  tidyr::unnest(cols = c(stats))  %>% 
  select(dw_mobilerecording_id, ends_with("mean"), ends_with("sd"))

stats
```


### Stats preview

We calculated stiatistics like this (only first 16 rows showed here):

```{r echo=FALSE}
colnames(stats) %>%
  sort() %>% 
  head(16)
```

For example, for the variable *f1frequency_sma3nz* statistics:

```{r echo=FALSE}
stats %>% select(starts_with("f1frequency"))
```

## Labels transformation examples

Transformation functions are used to re-code states into new values, like 'healty/unhealty' instead of 'euthymia/depression'.

```{r}
visits_label_hu <- transform_label_healthy_unhealthy(visits, 'hamd_ymrs')
visits_label_custom <- transform_label_custom(visits, 'hamd_ymrs')
visits_label_cgi <- transform_label_cgi(visits, 'hamd_ymrs')
```
